Domain-Driven, Generative Data Model for BigPetStore

### 2014-07-23 bdcloud_paper/latex/paper.tex
Updated HMM equations

A customer's purchases in each transaction are modeled using a layered Hidden Markov Model (HMM) (Figure~\ref{fig:shopping_cart_sim}).  The HMM has three types of states: the start state, the purchase states, and the stop state. There are an infinite number of purchase states, identified by the exhaustion times of the customer's item categories, the transaction time, and the number of items purchased in the current transaction. The observables of the states correspond to the item categories.  Each item category $i$ is assigned a weight $w_i$ according to the amount of time between the transaction time $T_t$ and $T_{E,i}$ when each item category will be exhausted (based on the usage simulations):

\begin{align*}
&w_i = - \beta^{-1}_c \exp(-\beta^{-1}_c (T_{E, i}- T_T)) \\
&\beta^{-1}_c \sim U(a,b)
\end{align*}

@@

\textcolor{red}{reorder so that probilities describe habits}

For example, if a customer tends to buy the same brand of dog food repeatedly, the outgoing edges to all products of the same brand as the product in the current state will be higher than those for other brands' products. If the customer does not care about flavor, the outgoing edges to all products from the current brand will be equal, otherwise the edge weights can be adjusted to prefer a specific combination of brand and flavor.  Likewise, if a customer cares more about flavor or cost, the edge weights will be assigned appropriately.

To simulate an item purchase, the Markov model is transitioned forward by one state.  The current states of the Markov models are kept between purchases.

After an item is chosen, the item category's exhaustion time is updated by propagating the item category usage SDE described in Section~\ref{sec:transaction-times}. \textcolor{red}{state transition, stop state}

The HMM can be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.

>>>>

A customer's purchases in each transaction are modeled using a layered Hidden Markov Model (HMM) (Figure~\ref{fig:shopping_cart_sim}).  The HMM has three types of states: the start state, the purchase states, and the stop state. There are an infinite number of purchase states, parameterized by the exhaustion times of the customer's item categories and the transaction time.

The observables of the states correspond to the item categories.  Each item category $i$ is assigned a weight $w_i$ according to the amount of time between the transaction time $T_t$ and $T_{E,i}$ when each item category will be exhausted (based on the usage simulations):

\begin{align} \label{eq:category-weights}
&w_i = - \beta^{-1}_c \exp(-\beta^{-1}_c (T_{E, i}- T_T)) \\
&\beta^{-1}_c \sim U(a,b) \nonumber
\end{align}

@@

%\textcolor{red}{reorder so that probilities describe habits}

%For example, if a customer tends to buy the same brand of dog food repeatedly, the outgoing edges to all products of the same brand as the product in the current state will be higher than those for other brands' products. If the customer does not care about flavor, the outgoing edges to all products from the current brand will be equal, otherwise the edge weights can be adjusted to prefer a specific combination of brand and flavor.  Likewise, if a customer cares more about flavor or cost, the edge weights will be assigned appropriately.

To simulate an item purchase, the chosen Markov model is transitioned forward by one state. The current states of the Markov models are kept between purchases.  After an item is chosen, the item category's exhaustion time is updated by propagating the item category usage SDE described in Section~\ref{sec:transaction-times}, resulting in an implicit creation of the next possible purchase state in the HMM.

After each purchase, the HMM's current state is transitioned to either the new purchase state or the stop state.  The probability of choosing the next state is

\begin{align*}
&Pr(X_{n+1}=\text{stop}|X_n) = p(\text{stop})\\
&Pr(X_{n+1}=\text{purchase}|X_n) = 1 - p(\text{stop})
\end{align*}

where $p(\text{stop})$ is the probability of choosing the stop state and is given by

\begin{equation*}
 p(\text{stop}) = \frac{w_{\text{stop}}}{w_{\text{stop}} + \sum_i w_i}
\end{equation*}

where $w_{\text{stop}}$ is the weight of the stop state and is a simulation parameter and the weight $w_i$ of each item category $i$ was given earlier in Eq.~\ref{eq:category-weights}.

%The HMM can be extended to consider additional factors such as the amount of money spent so far in the transaction and the customer's spending habits.
